Technical Report

### 2018-03-21 fully-abstract-comp/technical-report.tex
Remove old value relation.
fix seal offset bug in implementation of call

%\lau{16-10-2017: Fill this paragraph with $\scall{\offpc,\offsigma}{}{}$ implementation, i.e., the instructions $\scall[1]{\offpc,\offsigma}{r_1}{r_2} \dots \scall[\calllen]{\offpc,\offsigma}{r_1}{r_2}$ is equal to. Add convenience notation like $\update{\mem.a}{\scall{\offpc,\offsigma}{r_1}{r_2}} = \dots$}
\lau{16-10-2017: Implementation of $\scall{\offpc,\offsigma}{}{}$ may need registers for temporary registers - we should update the semantics accordingly.}
\[
  \begin{array}{l}
    \text{// push 42 on the stack (so it is non-empty).}\\
    \tmove{\rtmp{1}}{42}\\
    \tstore{\rstk}{\rtmp{1}}\\
    \tcca{\rstk}{(-1)}\\
    \text{// split the stack at its current address - rstk done.}\\
    \tgeta{\rtmp{1}}{\rstk}\\
    \tsplit{\rstk}{\rretd}{\rstk}{\rtmp{1}}\\
    \text{// load the seal for the return pointer through the pc capability.}\\
    \tmove{\rtmp{1}}{\pcreg}\\
    \tcca{\rtmp{1}}{\offpc} \text{ // $\offpc$ is the offset from pc to the location of the seal for this code}\\
    \tload{\rtmp{1}}{\rtmp{1}}\\
    \tcca{\rtmp{1}}{\offsigma}\text{ // $\offsigma$ is the offset for the seal used by this call.}\\
    \text{// seal the used stack frame as the data part of the return pointer pair.}\\
    \tcseal{\rretd}{\rtmp{1}}\\
    \text{// obtain the code part of the return pointer pair and seal it too.}\\
    \tmove{\rretc}{\pcreg}\\
    \tcca{\rretc}{5} \text{ //magic number is offset to return code}\\
    \tcseal{\rretc}{\rtmp{1}}\\
    \text{// now clear temporary register and jump to the adversary.}\\
    \tmove{\rtmp{1}}{0}\\
    \txjmp{r_1}{r_2}\\
    \text{// the following is the return code}\\
    \text{// check that the stack pointer is the same we handed out.}\\
    \tgetb{\rtmp{1}}{\rstk}\\
    \tminus{\rtmp{1}}{t_{t1}}{\stkb} \text{ //$\stkb$ is the stack base constant} \\
    \tmove{\rtmp{2}}{\pcreg}\\
    \tcca{\rtmp{2}}{5} \text{ //magic number is the offset to fail} \\
    \tjnz{\rtmp{2}}{\rtmp{1}} \\
    \tcca{\rtmp{2}}{1} \text{ //magic number is the offset to after fail} \\
    \tjmp{\rtmp{2}} \\
    \tfail \\
    \text{// join our stored private stack frame with the rest of the stack (this also finishes the stack pointer check).}\\
    \tsplice{\rstk}{\rstk}{\rdata} \\
    \text{// pop the magic number 42}\\
    \tcca{\rstk}{1}\\
    \text{// clear temporary registers used}\\
    \tmove{\rtmp{2}}{0}\\
    \text{// continue program after invocation.}
\]
\dominique{Return code could be simplified if we had jz in addition to jnz.}
% \dominique{It is a bit annoying that we store the stack end address before invocation if we do not use it.
%   I understand this is for ensuring that our private stack frame is not empty, which is important.
%   But why don't we just push a constant then?
%   Also: perhaps we could only do it if the current private stack frame is empty to begin with?
% \dominique{4-12-2017: update: Lau executed the first suggestion (push a constant instead), and convinced me that the last suggestion didn't make sense (only push if the current private stack is empty), because it is then not clear whether or not to pop in the return code.}
The call code does the following:
\begin{itemize}
\item Store 42 (it could be anything) to the stack (this ensures the stack is non-empty), and decrement the pointer according to convention.
\item Get the current address of the stack pointer and split according to it.
\item Retrieve the seal of the program.
\item cca the seal, so the seal to be used is active 
\item seal our private part of the stack capability.
\item Move the pc out of the pc register, adjust it to point to the first address of the return code, and seal it.
\item clear the temporary register.
\item cross jump to the two specified registers.
\item Upon return:
  \begin{itemize}
  \item get the base of the stack and check that it matches up with the global base of the stack.If not, fail.
  \item splice the returned stack pointer and the stack pointer for our private stack.
  \item adjust stack pointer to first empty address (the address with the end address is considered free)  
  \item clear the temporary registers. (Note that $\rtmp{1}$ is not cleared with 0 as it already contains 0 after the $\stkb$ check. If it didn't contain 0, then the execution would have failed.)

@@

\begin{definition}
  \label{def:points-to-call}
  A configuration {\sourcecolor $\Phi = (\ms,\reg,\_,\_)$} is said to point to instruction $\scall{o}{r_1}{r_2}$ in $A$ iff
  \[
    \reg(\pcreg) = ((\_,\_),\_,\eaddr,\aaddr) \tand [\aaddr,\aaddr+\calllen-1] \subseteq A
  \]
and
\[
  \aaddr+\calllen - 1 \leq \eaddr
\]
and
\[
  \callCond{\Phi,r_1,r_2,\aaddr}
\]
\end{definition}
We want to make a conditional full-abstraction statement. The condition should express two things (1) the compilation so far has been reasonable, so we do not have to setup protection against things any reasonable compiler wouldn't do. (2) the compilation so far has taken care of certain checks that we do not have information to place correctly at this point of the compilation.
\begin{definition}
  \label{def:check-stack-addr-before-call}
  A configuration $\src{\Phi}$ is said to check $\stkb$ before calls in $\ta$ for $n$ steps iff\\
  for all $i \leq n$ if
  \[
    \src{\Phi} \nstep[i] \src{\Phi'}
  \]
  and $\src{\Phi'}$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ in $\ta$ for some $\src{r_1}$ and $\src{r_2}$
  % and there exists no $j<i$ s.t.
  % \[
  %   \src{\Phi} \nstep[j] \src{\Phi''}
  % \]
  % where $\src{\Phi''}$ points to $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ in $A$ for some $\src{r_1'}$ and $\src{r_2'}$,
  then
  \[
    \src{\Phi'}(\src{r_\stk}) = \src{\stkptr{\_,\stkb,\_,\_}}
  \]
  and
  \[
    r_1' \neq \rtmp{1}
  \]
  % and $\src{\Phi'}$ has reasonable calls in $A$ for $n-i$ step.
\end{definition}

>>>>

%\lau{16-10-2017: Fill this paragraph with $\scall{\offpc,\offsigma}{}{}$ implementation, i.e., the instructions $\scall[1]{\offpc,\offsigma}{r_1}{r_2} \dots \scall[\calllen]{\offpc,\offsigma}{r_1}{r_2}$ is equal to. Add convenience notation like $\update{\mem.a}{\scall{\offpc,\offsigma}{r_1}{r_2}} = \dots$}
\lau{16-10-2017: Implementation of $\scall{\offpc,\offsigma}{}{}$ may need registers for temporary registers - we should update the semantics accordingly.}
\[
  \begin{array}{l}
    \text{// push 42 on the stack (so it is non-empty).}\\
    \tmove{\rtmp{1}}{42}\\
    \tstore{\rstk}{\rtmp{1}}\\
    \tcca{\rstk}{(-1)}\\
    \text{// split the stack at its current address - rstk done.}\\
    \tgeta{\rtmp{1}}{\rstk}\\
    \tsplit{\rstk}{\rretd}{\rstk}{\rtmp{1}}\\
    \text{// load the seal for the return pointer through the pc capability.}\\
    \tmove{\rtmp{1}}{\pcreg}\\
    \tcca{\rtmp{1}}{(\offpc - 5)} \text{ // $\offpc$ is the offset from pc to the location of the seal for this code. \footnote{-5 is the offset from the previous instruction where the $\pcreg$ was moved to $\rtmp{1}$ to the first instruction of the call implementation.}}\\
    \tload{\rtmp{1}}{\rtmp{1}}\\

    \tcca{\rtmp{1}}{\offsigma}\text{ // $\offsigma$ is the offset for the seal used by this call.}\\
    \text{// seal the used stack frame as the data part of the return pointer pair.}\\
    \tcseal{\rretd}{\rtmp{1}}\\
    \text{// obtain the code part of the return pointer pair and seal it too.}\\
    \tmove{\rretc}{\pcreg}\\
    \tcca{\rretc}{5} \text{ //magic number is offset to return code}\\
    \tcseal{\rretc}{\rtmp{1}}\\
    \text{// now clear temporary register and jump to the adversary.}\\
    \tmove{\rtmp{1}}{0}\\
    \txjmp{r_1}{r_2}\\
    \text{// the following is the return code}\\
    \text{// check that the stack pointer is the same we handed out.}\\
    \tgetb{\rtmp{1}}{\rstk}\\
    \tminus{\rtmp{1}}{t_{t1}}{\stkb} \text{ //$\stkb$ is the stack base constant} \\
    \tmove{\rtmp{2}}{\pcreg}\\
    \tcca{\rtmp{2}}{5} \text{ //magic number is the offset to fail} \\
    \tjnz{\rtmp{2}}{\rtmp{1}} \\
    \tcca{\rtmp{2}}{1} \text{ //magic number is the offset to after fail} \\
    \tjmp{\rtmp{2}} \\
    \tfail \\
    \text{// join our stored private stack frame with the rest of the stack (this also finishes the stack pointer check).}\\
    \tsplice{\rstk}{\rstk}{\rdata} \\
    \text{// pop the magic number 42}\\
    \tcca{\rstk}{1}\\
    \text{// clear temporary registers used}\\
    \tmove{\rtmp{2}}{0}\\
    \text{// continue program after invocation.}
  \end{array}
\]
\dominique{Return code could be simplified if we had jz in addition to jnz.}
% \dominique{It is a bit annoying that we store the stack end address before invocation if we do not use it.
%   I understand this is for ensuring that our private stack frame is not empty, which is important.
%   But why don't we just push a constant then?
%   Also: perhaps we could only do it if the current private stack frame is empty to begin with?
% }
% \dominique{4-12-2017: update: Lau executed the first suggestion (push a constant instead), and convinced me that the last suggestion didn't make sense (only push if the current private stack is empty), because it is then not clear whether or not to pop in the return code.}
The call code does the following:
\begin{itemize}
\item Store 42 (it could be anything) to the stack (this ensures the stack is non-empty), and decrement the pointer according to convention.
\item Get the current address of the stack pointer and split according to it.
\item Retrieve the seal of the program.
\item cca the seal, so the seal to be used is active 
\item seal our private part of the stack capability.
\item Move the pc out of the pc register, adjust it to point to the first address of the return code, and seal it.
\item clear the temporary register.
\item cross jump to the two specified registers.
\item Upon return:
  \begin{itemize}
  \item get the base of the stack and check that it matches up with the global base of the stack.If not, fail.
  \item splice the returned stack pointer and the stack pointer for our private stack.
  \item adjust stack pointer to first empty address (the address with the end address is considered free)  
  \item clear the temporary registers. (Note that $\rtmp{1}$ is not cleared with 0 as it already contains 0 after the $\stkb$ check. If it didn't contain 0, then the execution would have failed.)

@@

\begin{definition}
  \label{def:points-to-call}
  A configuration {\sourcecolor $\Phi = (\ms,\reg,\_,\_)$} is said to point to instruction $\scall{o}{r_1}{r_2}$ in $A$ iff
  \[
    \reg(\pcreg) = ((\_,\_),\_,\eaddr,\aaddr) \tand [\aaddr,\aaddr+\calllen-1] \subseteq A
  \]
and
\[
  \aaddr+\calllen - 1 \leq \eaddr
\]
and
\[
  \callCond{\Phi,r_1,r_2,\aaddr}
\]
\end{definition}
We want to make a conditional full-abstraction statement. The condition should express two things (1) the compilation so far has been reasonable, so we do not have to setup protection against things any reasonable compiler wouldn't do. (2) the compilation so far has taken care of certain checks that we do not have information to place correctly at this point of the compilation.
\begin{definition}
  \label{def:check-stack-addr-before-call}
  A configuration $\src{\Phi}$ is said to check $\stkb$ before calls in $\ta$ for $n$ steps iff\\
  for all $i \leq n$ if
  \[
    \src{\Phi} \nstep[i] \src{\Phi'}
  \]
  and $\src{\Phi'}$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ in $\ta$ for some $\src{r_1}$ and $\src{r_2}$
  % and there exists no $j<i$ s.t.
  % \[
  %   \src{\Phi} \nstep[j] \src{\Phi''}
  % \]
  % where $\src{\Phi''}$ points to $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ in $A$ for some $\src{r_1'}$ and $\src{r_2'}$,
  then
  \[
    \src{\Phi'}(\src{r_\stk}) = \src{\stkptr{\_,\stkb,\_,\_}}
  \]
  and
  \[
    r_1' \neq \rtmp{1}
  \]
  \lau{maybe the last bit should have its own definition.}
  % and $\src{\Phi'}$ has reasonable calls in $A$ for $n-i$ step.
\end{definition}
